### 2019-09-01 paper/paper.tex
Finished my first pass at editing.

    $\lambda$ & 0.39968             & \si{\radian} & Steer axis tilt \\

@@

    $\alpha_P$ & 0.18618 & \si{\radian} & Angle about $Y$ between $X$ and $A_P$ \\

@@

    $\alpha_H$ & 0.36995 & \si{\radian} & Angle about $Y$ between $X$ and $A_H$ \\

@@

    $\alpha_D$ &  1.1722 & \si{\radian} & Angle about $Y$ between $X$ and $A_D$ \\

@@

\section{RESULTS}
%
We discover four bicycles for different design speeds (3, 5, 7, and
9~\si{\meter\per\second}) that have an optimally low HQM, see
Table~\ref{tab:hqm}, and satisfy all constraints and parameter bounds. The
optimal parameter values for these four bicycles are presented in
Table~\ref{tab:optimal-values}. We belive these bicycles to be reasoably
physically realizable. The pictorial representation of the bicycles are
presented in Figure~\ref{fig:optimal-geometries}.
%
\begin{table}
  \caption{Peak HQM values for the reference bicycle and the optimal bicycles
    at each design speed.}
  \label{tab:hqm}
  \centering
  \begin{tabular}{llll}
    \toprule
    Speed [m/s] & Reference Peak HQM & Optimal Peak HQM & Percent Improvement \\
    \midrule
    3 & 13.075 & 2.012 & 85\% \\
    5 & 4.521  & 0.012 & 100\% \\
    7 & 3.043  & 0.022 & 99\% \\
    9 & 2.338  & 0.839 & 64\% \\
    \bottomrule
  \end{tabular}
\end{table}
%
\begin{table}
  \caption{Optimal principal parameter values for each design speed. Values
    with an asterisk are at a bound.}
  \label{tab:optimal-values}
  \centering
  \begin{tabular}{lSSSS}
    \toprule
    $v$ & 3~\si{\meter\per\second} & 5~\si{\meter\per\second} & 7~\si{\meter\per\second} & 9~\si{\meter\per\second} \\
    \midrule
    $c$       & 0.688  & -0.005 & 0.001  & -0.484 \\
    $w$       & 2.866  & 0.847  & 1.157  & 5.557  \\
    $\lambda$ & -0.213 & 0.271  & -0.028 & 0.239  \\
    \midrule
    $m_D$ & 3.22   &  11.53 & 10.32  & 7.02 \\
    $x_D$ & 0.958  &  0.287 &  0.427 & 1.020 \\
    $z_D$ & -2.605 & -2.719 & -2.976 & -4.065 \\
    $\alpha_D$ & 0.663  & 0.915  & 1.123  & 1.077 \\
    % Should aa be larger than bb?
    $k_{Daa}$  & 1.449  & 1.218  &  1.324 & 0.861 \\
    $k_{Dbb}$  & 0.724  & 1.387  &  1.290 & 2.014 \\
    $k_{Dyy}$  & 1.471  & 1.559  &  1.555 & 2.031 \\
    \midrule
    $m_H$  & 0.25*  & 0.25* &    0.49 &   3.54 \\
    $x_H$     & 2.356   & 0.532 &    1.145 &    5.615 \\
    $z_H$     & -1.298  & -0.781 &   -1.340 &   -1.776 \\
    $\alpha_H$ & 0.572  & -1.265 & 1.571  & 0.238 \\
    $k_{Haa}$ & 0.186  & 0.0491 &   0.000* &       2.846 \\
    $k_{Hbb}$ & 1.259   & 1.145 &     2.006 &    0.168 \\
    $k_{Hyy}$ & 0.636  & 0.383 &    0.818 &   0.669 \\
    \midrule
    $m_F$     & 1.62  & 4.40  & 2.64  & 5.51  \\
    $r_F$     & 0.710 & 0.252 & 0.127* & 2.063 \\
    $k_{Faa}$ & 0.355 & 0.126 & 0.064 & 1.031 \\
    $k_{Fyy}$ & 0.710 & 0.252 & 0.127 & 2.063 \\
    \midrule
    $x_P$      & 0.765  & 0.276  &  0.526 & 1.079 \\
    $z_P$      & -3.194 & -0.453 & -0.586 &-2.662 \\
    $\alpha_P$ & -0.661 & 1.150  & -0.836 & 1.558 \\
    \midrule
    $m_R$   & 12.63  & 8.36  &  6.31 & 2.25 \\
    $r_R$   & 0.958  & 0.506 & 0.146 & 3.027 \\
    $k_{Raa}$ & 0.479  & 0.253 & 0.073 & 1.514 \\
    $k_{Ryy}$ & 0.958  & 0.506 & 0.146 & 3.027 \\
    \bottomrule
  \end{tabular}
\end{table}
%
\begin{figure}
  \centering
  \includegraphics[width=6in]{../figures/optimal-geometries.png}
  \label{fig:optimal-geometries}
  \caption{Depictions of the bicycle geometry and geometric representations of
    the inertial quantities for the reference bicycle and four optimal solutions
    at 3, 5, 7, and 9 m/s. Five rigid bodies are shown for each bicycle: front
    wheel (orange), rear wheel (purple), rear frame (blue), front frame
    (green), and person (red). The solid black lines represent the essential
    bicycle geometry. The dotted black line represents the steer axis. The
    solid colored curves represent the contours of solid ellipsoids with
    equivalent inertia as the principal inertia of the associated rigid body.
    The dotted colored lines represent the extents of the centroidal radii of
    gyration of each rigid body.}
\end{figure}

The bicycle optimized for 3~\si{\meter\per\second} is about three times larger
than the reference bicycle. The wheelbase is about 3~\si{\meter} and it has a
relatively large postive trail (0.7~\si{\meter}). The person is rearward on the
bike and over 3~\si{\meter} above the ground. The mass of the front frame is
the minimum posible value and the mass of the rear frame is also small.

The bicycle optimized for 5~\si{\meter\per\second} has a similar geometric
scale as the reference bicycle except that the front wheel is smaller and the
rear wheel is larger. The person is located low to the ground and tipped back,
like one might see on a recumbent bicycle.  This significantly increases the
vehicles yaw moment of inertia. The trail is miminal, but suprisingly slight
negative (5~\si{\milli\meter}). The steer axis tilt is slightly shallower than
the reference bicycle. The mass of the rear frame is similar to the reference
bicycle but the front frame is at the bound and much lower. The minor principal
axis of the front frame is almost aligned with the steer axis.

The bicycle optimized for 5~\si{\meter\per\second} is one of the three that is
geometrically smaller than the reference bicycle. The wheelbase is slightly
larger but the wheels are approximately half the diameter. The rider is low to
the ground and tipped forward. There is effiectively no trail, the steer axis
is vertical, and the (very) minimal principal axis is aligned with the steer
axis making the steering inertia-less. The roll moment of inertia is larger.

The bicycle optimized for 5~\si{\meter\per\second} is much larger than the
reference bicycle: wheelbase and rear wheel diameter of 6~\si{\meter}, front
wheel diameter of 4~\si{\meter}. The trail is large and negative, surely making
the bicycle open-loop unstable.

The parameters of the bicycles can be distilled into open loop eigenvalues for
a clearer understanding of the dynamics, shown in
Figure~\ref{fig:optimal-eigenvalues}. The 3 and 9 design speed have a simlar
eigenvalue pattern as well as the 5 and 7. The 7~\si{\meter\per\second} design
is self-stable at speeds greater than 1.2~\si{\meter\per\second} but the other
three designs are not self-stable at any travel speed from
\SIrange{0}{10}{\meter\per\second}. Both the 5 and 7 designs have a weave mode
that increases in frequency rapidly with speed. The weave frequency of the 3
m/s design is higher than the reference design and present at all speeds.
%
\begin{figure}
  \centering
  \includegraphics[width=6in]{../figures/optimal-eigenvalues.png}
  \label{fig:optimal-eigenvalues}
  \caption{Real (solid) and imaginary (dashed) parts of the open-loop eigenvlues plotted versus
    travel speed. The vertical dashed black line indicates the design speed.}
\end{figure}

@@

The bicycles are reasonably physically realizable but some additional
constraints could be introduced to further improve this. In the
5~\si{\meter\per\second} design the person's body overlaps the wheels, which
would be almost impossible to realize. A constraint that ensures the torso
cannot occupy the same space as the wheels would solve this. The masses of the
front and rear frames are low with respect to the geometrical spread a steel
space frame would have to occupy. The frames also need to be able to span the
space between the respecitve wheel, mass center of the frame, and the steer
axis and for the rear frame span to the rider's support locations. We currently
only expliclty deal with spanning the space between the mass center of the
frame and the respective wheel. Another more restricting appraoch would be to
specify a generic geometric structure for the front and rear frames. If done,
maximal stress and deflections of the frames under load could also be
minimized.

Each optimizal solution requires approximately 12 hours of computation time on
a high end consumer desktop machine running on a single thread. The number of
iterations are typically in the hundreds of thousands for convergence. There
are many avenues for speeding this up that could likely result in solutions in
one or two hours on the same machine but this is still slow to be particualrly
pleasant to utlizie in the process of designing a bicycle. Furthermore, the
time required to translate the resulting parameters in to an actual bicycle
design complete with structural destails is extremely time consuming.
Never-the-less, the method shows promise for optimizing an entire vehicle for
optimal dynamics. This method can be applied to a whole host of human operated
vehicles opening up many new designs, but as with any optimization it only
captures a very small set of the variables that a designer has to take into
account for a vehicle.

>>>>

    $\lambda$ & 0.400             & \si{\radian} & Steer axis tilt \\

@@

    $\alpha_P$ & 0.186 & \si{\radian} & Angle about $Y$ between $X$ and $A_P$ \\

@@

    $\alpha_H$ & 0.370 & \si{\radian} & Angle about $Y$ between $X$ and $A_H$ \\

@@

    $\alpha_D$ &  1.172 & \si{\radian} & Angle about $Y$ between $X$ and $A_D$ \\

@@

\section{RESULTS}
%
We discover four bicycles for different design speeds (3, 5, 7, and
9~\si{\meter\per\second}) that have an optimally low HQM, see
Table~\ref{tab:hqm}, and satisfy all constraints and parameter bounds. The
optimal parameter values for these four bicycles are presented in
Table~\ref{tab:optimal-values}. We believe these bicycles to be reasonably
physically realizable. The pictorial representation of the bicycles are
presented in Figure~\ref{fig:optimal-geometries}.
%
\begin{table}
  \caption{Peak HQM values for the reference bicycle and the optimal bicycles
    at each design speed.}
  \label{tab:hqm}
  \centering
  \begin{tabular}{llll}
    \toprule
    Speed [m/s] & Reference Peak HQM & Optimal Peak HQM & Percent Improvement \\
    \midrule
    3 & 13.075 & 2.012 & 85\% \\
    5 & 4.521  & 0.012 & 100\% \\
    7 & 3.043  & 0.022 & 99\% \\
    9 & 2.338  & 0.839 & 64\% \\
    \bottomrule
  \end{tabular}
\end{table}
%
\begin{table}
  \caption{Optimal principal parameter values for each design speed. Values
    with an asterisk are at a parameter bound.}
  \label{tab:optimal-values}
  \centering
  \begin{tabular}{lSSSS}
    \toprule
    $v$ & 3~\si{\meter\per\second} & 5~\si{\meter\per\second} & 7~\si{\meter\per\second} & 9~\si{\meter\per\second} \\
    \midrule
    $c$       & 0.688  & -0.005 & 0.001  & -0.484 \\
    $w$       & 2.866  & 0.847  & 1.157  & 5.557  \\
    $\lambda$ & -0.213 & 0.271  & -0.028 & 0.239  \\
    \midrule
    $m_D$ & 3.22   &  11.53 & 10.32  & 7.02 \\
    $x_D$ & 0.958  &  0.287 &  0.427 & 1.020 \\
    $z_D$ & -2.605 & -2.719 & -2.976 & -4.065 \\
    $\alpha_D$ & 0.663  & 0.915  & 1.123  & 1.077 \\
    % Should aa be larger than bb?
    $k_{Daa}$  & 1.449  & 1.218  &  1.324 & 0.861 \\
    $k_{Dbb}$  & 0.724  & 1.387  &  1.290 & 2.014 \\
    $k_{Dyy}$  & 1.471  & 1.559  &  1.555 & 2.031 \\
    \midrule
    $m_H$  & 0.25*  & 0.25* &    0.49 &   3.54 \\
    $x_H$     & 2.356   & 0.532 &    1.145 &    5.615 \\
    $z_H$     & -1.298  & -0.781 &   -1.340 &   -1.776 \\
    $\alpha_H$ & 0.572  & -1.265 & 1.571  & 0.238 \\
    $k_{Haa}$ & 0.186  & 0.0491 &   0.000* &       2.846 \\
    $k_{Hbb}$ & 1.259   & 1.145 &     2.006 &    0.168 \\
    $k_{Hyy}$ & 0.636  & 0.383 &    0.818 &   0.669 \\
    \midrule
    $m_F$     & 1.62  & 4.40  & 2.64  & 5.51  \\
    $r_F$     & 0.710 & 0.252 & 0.127* & 2.063 \\
    $k_{Faa}$ & 0.355 & 0.126 & 0.064 & 1.031 \\
    $k_{Fyy}$ & 0.710 & 0.252 & 0.127 & 2.063 \\
    \midrule
    $x_P$      & 0.765  & 0.276  &  0.526 & 1.079 \\
    $z_P$      & -3.194 & -0.453 & -0.586 &-2.662 \\
    $\alpha_P$ & -0.661 & 1.150  & -0.836 & 1.558 \\
    \midrule
    $m_R$   & 12.63  & 8.36  &  6.31 & 2.25 \\
    $r_R$   & 0.958  & 0.506 & 0.146 & 3.027 \\
    $k_{Raa}$ & 0.479  & 0.253 & 0.073 & 1.514 \\
    $k_{Ryy}$ & 0.958  & 0.506 & 0.146 & 3.027 \\
    \bottomrule
  \end{tabular}
\end{table}
%
\begin{figure}
  \centering
  \includegraphics[width=6in]{../figures/optimal-geometries.png}
  \label{fig:optimal-geometries}
  \caption{Depictions of the bicycle geometry and geometric representations of
    the inertial quantities for the reference bicycle and four optimal
    solutions at 3, 5, 7, and 9~\si{\meter\per\second}. Five rigid bodies are
    shown for each bicycle: front wheel (orange), rear wheel (purple), rear
    frame (blue), front frame (green), and person (red). The solid black lines
    represent the essential bicycle geometry. The dashed black line represents
    the steer axis. The dashed colored lines represent the extents of the
    centroidal radii of gyration of each rigid body and the red diamond shows
    the physical extents of the persons geometry.}
\end{figure}

The bicycle optimized for 3~\si{\meter\per\second} is about three times larger
than the reference bicycle. The wheelbase is about 3~\si{\meter} and it has a
relatively large positive trail (0.7~\si{\meter}). The person is rearward on
the bike and over 3~\si{\meter} above the ground. The mass of the front frame
is the minimum possible value and the mass of the rear frame is also small.

The bicycle optimized for 5~\si{\meter\per\second} has a similar geometric
scale as the reference bicycle except that the front wheel is smaller and the
rear wheel is larger. The person is located low to the ground and tipped back,
like one might see on a recumbent bicycle. This significantly increases the
vehicles yaw moment of inertia. The trail is minimal, but surprisingly slightly
negative (5~\si{\milli\meter}). The steer axis tilt is slightly shallower than
the reference bicycle. The mass of the rear frame is similar to the reference
bicycle but the front frame is at the bound and much lower. The minor principal
axis of the front frame is almost aligned with the steer axis.

The bicycle optimized for 7~\si{\meter\per\second} is one of the three that is
geometrically smaller than the reference bicycle. The wheelbase is slightly
larger but the wheels are approximately half the diameter. The rider is low to
the ground and tipped forward. There is effectively no trail, the steer axis is
vertical, and the (very) minimal principal axis is aligned with the steer axis
making the steering effectively inertia-less. The roll moment of inertia is
larger than the reference bicycle.

The bicycle optimized for 9~\si{\meter\per\second} is much larger than the
reference bicycle: wheelbase and rear wheel diameter of 6~\si{\meter}, front
wheel diameter of 4~\si{\meter}. The trail is large and negative, surely making
the bicycle open-loop unstable. The person is in a recumbent position and very
rearward. The mass centers of the person, front and rear frames are very close
to the respective wheel mass centers. The minor principal axis of the front
frame is aligned with the steer axis, reducing the steer inertia. The masses of
the two frames do not seem to be sufficient for such a large steel space frame.

The parameters of the bicycles can also be distilled into open loop eigenvalues
for a clearer picture of the dynamics, shown in
Figure~\ref{fig:optimal-eigenvalues}. The 3 and 9~\si{\meter\per\second}
designs have a similar eigenvalue pattern as does the 5 and
7~\si{\meter\per\second} designs. The 7~\si{\meter\per\second} design is
self-stable at speeds greater than 1.2~\si{\meter\per\second} but the other
three designs are not self-stable at any travel speed from
\SIrange{0}{10}{\meter\per\second}. Both the 5 and 7~\si{\meter\per\second}
designs have a weave mode that increases in frequency rapidly with speed. The
weave frequency of the 3~\si{\meter\per\second} design is higher than the
reference design and present at all speeds.
%
\begin{figure}
  \centering
  \includegraphics[width=6in]{../figures/optimal-eigenvalues.png}
  \label{fig:optimal-eigenvalues}
  \caption{Real (solid) and imaginary (dashed) parts of the open-loop
    eigenvalues plotted versus travel speed. The vertical dashed black line
    indicates the design speed.}
\end{figure}

@@

The bicycles are reasonably physically realizable but some additional
constraints could be introduced to further improve this. In the
5~\si{\meter\per\second} design the person's body overlaps the wheels, which
would be almost impossible to realize. A constraint that ensures the torso
cannot occupy the same space as the wheels would solve this. The masses of the
front and rear frames are low with respect to the geometrical spread a steel
space frame would have to occupy. The frames also need to be able to span the
space between the respective wheel, mass center of the frame, and the steer
axis and for the rear frame span to the rider's support locations. We currently
only explicitly deal with spanning the space between the mass center of the
frame and the respective wheel. Another more restricting approach would be to
specify a generic geometric structure for the front and rear frames. If done,
maximal stress and deflections of the frames under load could also be
minimized.

Each optimized solution requires approximately 12 hours of computation time on
a high-end consumer desktop computer running the optimization procedure
sequentially. The number of iterations are typically in the hundreds of
thousands for convergence. There are many avenues for speeding this up that
could likely result in solutions in one or two hours on the same computer
hardware but this is still slow to be particularly convenient to utilize in the
process of designing a bicycle. Furthermore, the time required to translate the
resulting parameters into an actual bicycle design complete with structural
details is extremely time consuming.  Never-the-less, the method shows promise
for optimizing an entire vehicle for optimal dynamics. This method can be
applied to a whole host of human operated vehicles opening up many new designs,
but as with any optimization it only captures a very small set of the variables
that a designer has to take into account for a vehicle.
